Renewal and stability in populations structured by remaining years of life

### 2015-11-09 ThanoRepro/ThanoRepro.tex
restructuring lead-in

\begin{abstract}
I present a unisex model of population renewal for populations structured by
remaining years of life in both continuous and discrete form.
Various stable properties of this model are compared and related when viewed
by chronological versus thanatological age.\ackn{Research reported in this manuscript
was supported by the U.S.
National Institute On Aging of the National Institutes of Health under award numbers R01-AG011552 and R01-AG040245 and by the Spanish Ministry of Science and Innovation under award
number ECO2013-48326-C2-1-P. The content is solely the responsibility of the
author and does not necessarily represent the official views of the funding
agencies.}
\end{abstract}

\onehalfspacing
\section*{Introduction}
In this paper, I explore some formal demographic consequences of a particular redefinition of age. Instead of counting age as the time passed since birth, consider age as the amount of time left until death\footnote{This quantity is referred to as \textit{residual} lifetime
in the reliability literaure, or just time-to-death (TTD) elsewhere.}. The first
kind of age is called chronological age, and the second kind of age is called
thanatological age. Individuals in this case move in the same direction along
imaginary life lines, but the reference point for thanatological age is at
the end of the life line instead of at the beginning. For individuals, the timing of events in life is exactly identified using either of the two definitions of age, but
aggregate patterns differ. Demography is built upon a small set of empirical
regularities with respect to chronological age patterns. The models and methods
of demography are in varying degrees prefaced on such regularity. There is evidence that many demographic phenomena are best described in the
aggregatate as a function of time since birth, and others of time until death,
while others can be a function of both to some degree
\citep{riffe2015ttd,wolf2015disability}. This observation motivates
incorporating this aspect of time into formal demographic methods and models.
%Temporal perspectives and rescalings of the lifecourse are abundant in
% demography already. Such work has enriched the discipline by producing new
% insights, questions, and conundrums.
% Fertility and reproduction are plausibly a function of both age perspectives,
 % and the question of how to best partition them over time is not investigated here.

I describe the Lotka-Euler-Leslie renewal model when structured by
thanatological age. The renewal model relates
fertility, mortality, and population structure in a parsimonious form, revealing the long term consequences of a particular set of conditions. All diverse lines of contemporary demographic research are relatable to the renewal model, and by extention to stable population theory.
%Some might say that stable population
%theory is not relevant to everyday demography, since populations are always
% changing. Even if change precludes the notion of stability as a prediction, models of long-run consequnces still give a
%useful measure of the demographic force implied by a given set of rates.
%In this and other ways, stability is never far away. That population
% projections lead into a stable state means that some form of stability is implied by the routine practice of demography.
%Stability or even stationarity are imaginable for the world population of
% humans one day. Models of stationary populations are easily applied by analogy to
%closed or nearly-closed birth cohorts. Finally, stable population theory is
%relavant for modelling some non-human populations. 
%We have stated that many lifecourse transitions covary more with remaining
%lifetime than with life lived. In an even more general sense, by thinking about
% remaining lifetime as a particular class of time-to-transition, one may generalize the lifetable methods presented here into new area applications, such as education, partnering, the economic lifecycle, or migration.
To present the renewal model based on thanatological age is not necessarily a
suggestion for a better way to model or project population. Rather, the
thanatological renewal model represents an alternative and coherent way to
describe the process of population renewal by relating renewal to a death flow
rather than to a birth flow. I have described a limited case of this
perspective on renewal for the case of stationary populations
\citep{riffe2015force}, and here provide a broader treatment allowing for
growth in the population.
The chronological-age classified renewal model and the thanatological renewal model both describe the movement of population stocks over time, but from opposite vantage points. Specifying a model of thanatological renewal expands the toolbox available to approach complex
temporal population structure.

The next section of this article describes the lifetable method of transforming age-classified
data into thanatological age classes. I then derive the continuous version of
the thanatological renewal model, and some other stable properties. This is
followed by the matrix version of the model, which is explained alongside the
lifecycle graph.

\section*{A lifetable approximation of thanatological age}
%Direct information on the timing of demographic transitions with respect to
%death is available retrospectively from linked population registers or
%panel surveys with mortality follow-ups.
%Some surveys also ask respondents in various ways about their perceived future
% lifetime, which may underly those demographic transitions governed to some degree by volition. 
In general, to make a lifetable-based projective statement about the remaining
time distribution of a current stock of population, it is best to to account for population subgroups
with different mortality and to account for uncertainty about future changes in
mortality.
The present inquiry approximates thanatological age using lifetable-projected remaining lifetime,
which is intended to approximate remaining lifetime under the strong
assumptions of homogenous populations und fixed vital rates. 
%Chronological age-classified counts are redistributed over future death times
% according to a fixed mortality schedule.
%In general, a lifetable approach to transforming chronologically classified
%data into thanatologically classified data is sound if the given lifetable
%accurately describes the future mortality of the age-classified population, and
%if the chronological and thantological age patterns are independent.
Under these assumptions, each birth cohort undergoes the same
pattern of attrition. In a period setting, this means that we can distribute the
remaining lifespans of the population in each age class using the same
lifetable.
%This is the case we explore, and so the following formulas assume a single lifetable pattern.

@@

A convenient discretization of \eqref{eq:vaupel1} is to simply work with the
$d_a$ column of the lifetable (here with single ages):
\begin{equation}
\label{eq:dyadiscret}
f_{y, a} =  \frac{d_{a+y}+d_{a+y+1}}{L_a + L_{a+1}} \ec
\end{equation}
where $L_a$ is lifetable exposure.\footnote{Equation~\eqref{eq:dyadiscret}
will incur some error depending on the method used to derive $L_a$. To remove
this error, rescale over the $a$ margin so that each distribution conditioned on
$a$ sums to 1. Working with $\ell_a$ rather than $L_a$ eliminates error and
yields similar results, but is not necessarily better.} One produces in this way a probability density function over future death times for each chronological age,
$a$. The $f(y|a)$- weighted average of $y$ for a given $a$ returns the familiar
remaining life expectancy column of the lifetable, $e(a)$. If $P(a)$ is a
chronological age-classified population count, we can derive $P(y)$, a
thanatological age-classified population count using\footnote{An
different-looking but identical definition is found in
\citet{brouard1989mouvements}.}

@@

Lifetable-derived thanatological age structure is
probabilistic until after death, since it refers to the future according to an
assumed mortality pattern; in this case a period mortality pattern is used.
As such, Figure~\ref{fig:USrecomp} is \textit{projective} in nature, a
forward-looking glance at population attrition given the population stock and lifetable of a
particular moment, whereas Figure~\ref{fig:USdecomp} is \textit{reflective} in
nature, since a population's chronological age-structure is mostly the fruit of past
fertility, but also to a lesser degree migration and attrition. Any count
classified by chronological age can be reclassified in this way given a
pertinent lifetable. To
derive thanatological fertility rates under these assumptions, $\gamma(y)$, apply \eqref{eq:transform} to
chronological age-classified birth counts, $B(a)$, to get $B(y)$ and again to
exposure-to-risk, $E(a)$, to calcualte $E(y)$ and then divide in the usual way:

@@

\FloatBarrier
\section*{The thanatological renewal model}
The renewal model that follows uses thanatological fertility rates, $\gamma(y)$,
the form of which is suggested in Figure~\ref{fig:FySpaghetti}, and we refer to
a homogenous, closed, unisex population based on female vital rates. Assuming constant vital rates, the births for the present year are given by 
\begin{align}
B(t) =& \int _0^\infty P(y,t)\gamma(y) \dd y = \int _0^\infty
P(a,t)m(a) \dd a \ec \intertext{where $a$ indexes chronological age, $y$ indexes
thanatological age, $P(a), P(y)$ are population counts, and $\gamma(y), m(a)$ are
exact fertility probabilities (rates), best imagined
as the fertility rates of females born to females. The thanatological
(first) integral can be broken down back in terms of chronological age:} =&
\int_{y=0}^\infty \int_{a=0}^\infty \gamma(y)P(a,t)f(y|a) \dd a \dd y \ec
\intertext{where $f(y|a)$ is defined per equation~\eqref{eq:vaupel1}, so
$d(a)/\ell(0)$.
We can relate the present population to past births with $P(a,t) = B(t-a)\ell(a)$:}
\label{eq:ergostep1}
=& \int_{y=0}^\infty \int_{a=0}^\infty \gamma(y) B(t-a)d(a+y)\dd a \dd y \ep 
\intertext{Eventually strong ergodicity will assert itself, and $B(t)$ will
be related to $B(t-a)$ according to a constant factor $e^{ra}$, where $r$ is the familiar
intrinsic rate of growth:}
\label{eq:ergostep2}
=& \int_{y=0}^\infty \int_{a=0}^\infty \gamma(y) B(t)e^{-ra}d(a+y)\dd a \dd y
\ep
\intertext{Divide out $B(t)$ to get back a familiar-looking renewal equation:}
\label{eq:thanoren}
1 =& \int_{y=0}^\infty \int_{a=0}^\infty \gamma(y) d(a+y)e^{-ra}\dd a \dd y
\ep
\end{align}
Now compare this to Lotka's chronological formulation:
\begin{align}
\label{eq:lotka}
1 =& \int_{a=0}^\infty m(a)l(a)e^{-ra}\dd a \ec
\intertext{which can be expressed as well as}
  =& \int_{a=0}^\infty \int_{y=0}^\infty m(a)d(a+y)e^{-ra}\dd y \dd a \ec
  \label{eq:lotka2}
\end{align}
and note that \eqref{eq:lotka2} and \eqref{eq:thanoren} are really quite
similar, since $\int _{y=0}^\infty d(a+y)\dd a = l(a)$. For intuition, notice
that $l(a)$ is here decomposed into $d(a)$. Imagine a matrix of this
decomposition, where one margin is chronological age and the other margin is
thanatological age.
For the chronological case \eqref{eq:lotka}, we multiply chronological
age-specific fertility rates, $m(a)$, over the chronological age margin, and for the thanatological case we multiply $\gamma(y)$ over the thanatological age margin.

@@

\section*{Discussion}
The thanatological renewal model is valid, but this perspective may take some
time to gestate before accepting that it is also a sound description of how
populations renew. What should one imagine under the model of thanatological
population renewal? A useful mnemonic bases itself on Figures~\ref{fig:USdecomp}
and \ref{fig:USrecomp}. In the chronological age-structured model, new
generations appear at the bottom of the pyramid, and move up one rung per year.
All age-classes are subject to attrition, which is spread out over ages and not
readily visible in the pyramid. In the thanatological \textit{leaf}, each birth cohort increments to the population over the whole range of thanatological age
according to $d(a)$, as seen in \eqref{eq:thanoren}, becoming the
shaded layers seen in Figure~\ref{fig:USrecomp}. Each horizontal step is a death
cohort, and these move one step down the pyramid each year without any decrement
(indeed incrementing due to births) until reaching the very bottom. In short,
the locations of increment and decrement, and the direction of movement (when
so visualized) are all switched. The chronological and thanatological renewal
models are almost perfectly opposite descriptions of the same process.

@@

\section{Equivalence in stability}
\label{app:B}
This appendix contains a proof that the thanatological renewal equation implies
the same intrinsic growth rate, $r$, as the Lotka-Euler model if
the starting population is stable, given a particular set of thanatological
fertility rates.
This is identical to claiming that the right side of the renewal equation is
equal given the same $r$. Since there is a unique real solution, demonstrated in the previous appendix, then it is sufficient to
show that the two integral equations are equal under these conditions. The basic
relationship is:
\begin{align}
\int _{a=0}^\infty \ell(a)e^{-ra}f(a) \dd a &= \int _{y=0}^\infty \int
_{a=0}^\infty d(a+y)e^{-ra}\gamma (y) \dd a \dd y
\intertext{when}
\gamma(y) &= \frac{b^\star(y)}{c^\star(y)} \\
b^\star (y) &= \int_{a=0}^\infty b(a) \mu(a+y)\frac{\ell(a+y)}{\ell(a)} \dd a \\
c^\star (y) &= \int_{a=0}^\infty c(a) \mu(a+y)\frac{\ell(a+y)}{\ell(a)} \dd a
\intertext{Note that:}
c(a) &= \frac{\ell(a)e^{-ra}}{\int \ell(a) e^{-ra} \dd a} \\
b(a) &= f(a)c(a)
\end{align}Replacing $\gamma(y)$ with the full expanded
expression and plugging into the thanatological renewal equation gives:
\begin{align}
=& \int _{y=0}^\infty \int
_{a=0}^\infty d(a+y)e^{-ra} \frac{\frac{\int_{t=0}^\infty
f(t)\ell(t)e^{-rt}\mu(a+y)\frac{\ell(t+y)}{\ell(t)} \dd t}{\int \ell(t) e^{-rt} \dd
t}}{\frac{\int_{t=0}^\infty \ell(t)e^{-rt} \mu(t+y)\frac{\ell(t+y)}{\ell(t)} \dd t}{\int
\ell(t) e^{-rt} \dd t}} \dd a \dd y
\intertext{The denominator of $c(a)$ cancels out:}
=& \int _{y=0}^\infty \int
_{a=0}^\infty d(a+y)e^{-ra} \frac{
\int_{t=0}^\infty
f(t)\ell(t)e^{-rt}\mu(t+y)\frac{\ell(t+y)}{\ell(t)} \dd t}{
\int_{t=0}^\infty \ell(t)e^{-rt} \mu(t+y)\frac{\ell(t+y)}{\ell(t)} \dd t} \dd a \dd y
\intertext{Some $l(a)$'s also cancel out:}
=& \int _{y=0}^\infty \int
_{a=0}^\infty d(a+y)e^{-ra} \frac{
\int_{t=0}^\infty
f(t)e^{-rt}\mu(t+y)\ell(t+y) \dd t}{
\int_{t=0}^\infty e^{-rt} \mu(t+y)\ell(t+y) \dd t} \dd a \dd y
\intertext{Now $d(a+y) = \mu (a+y)\ell(a+y)$, so we can cancel the inner integral
with the denominator:}
=& \int _{y=0}^\infty 
\int_{a=0}^\infty
f(a)e^{-ra}\mu(a+y)\ell(a+y) \dd a \dd y
\intertext{Note also that $l(a) = \int _{=a}^\infty d(x) \dd x$ (present
livings are future deaths), which brings us back to the chronological
formulation:} =& \int_{a=0}^\infty e^{-ra}\ell(a)f(a)
\dd a
\end{align}
This finding essentially proves that in the stable population there is at least
one thanatological fertility rate schedule that will satisfy the constraints of
the lifespan distribution and $r$. However, it is not necessary to derive the
thanatological fertility distribution in the way prescribed here, and indeed an
infinite number of such fertility distributions would satisfy the same
constraints. Imagine these fertility rates cross-classified in
both age dimensions, where two marginal sums give the thanatological and
chronological rates, as they would be calculated over the
corresponding cross-classified exposures. The shape of this rate surface can be
shifted around in infinitely many ways that would preserve a given growth rate
and lifespan distribution, and the above proof shows one such mapping.

>>>>

\begin{abstract}
Models of age structured population renewal are the core of formal demography.
I derive a specification of the classic renewal equation that is structured by
thanatological age rather than by chronological age. I give both continuous
and discrete variants of the derived model, and relate these to the well-known
age structured model. In stability, the two kinds of age
structure are commensurable, and imply identical intrinsic growth rates.
I demonstrate approximate symmetry between chronological and
thanatological age structure in stability when subject to
intrinsic growth rates equal magnitude and opposite sign. This paper offers a
new perspective on the Lotka-Leslie renewal process.\ackn{Research reported in
this manuscript was supported by the U.S.
National Institute On Aging of the National Institutes of Health under award numbers R01-AG011552 and R01-AG040245. The content is solely the responsibility of the
author and does not necessarily represent the official views of the funding
agencies.}
\end{abstract}

\onehalfspacing
\section*{Introduction}

Demography is built upon a small set of empirical
regularities with respect to chronological age patterns. The models and methods
of demography are in varying degrees prefaced on such regularity. There is evidence that many demographic phenomena are best described in the
aggregatate as a function of time since birth, and others of time until death,
while others can be a function of both to some degree
\citep{riffe2015ttd,wolf2015disability}. This observation motivates
incorporating this aspect time-to-death into formal demographic methods and
models.

In this paper, I explore some formal demographic consequences of a particular redefinition of age. Instead of counting age as the time passed since birth, consider age as the amount of time left until death. The first
kind of age is called chronological age, and the second kind of age is called
thanatological age.\footnote{This quantity is referred to as \textit{residual} lifetime
in the reliability literaure, or just time-to-death (TTD) elsewhere.} Individuals in this case move in the same direction along
imaginary life lines, but the reference point for thanatological age is at
the end of the life line instead of at the beginning. For individuals, the timing of events in life is exactly identified using either of the two definitions of age, but
aggregate patterns differ. 
%Temporal perspectives and rescalings of the lifecourse are abundant in
% demography already. Such work has enriched the discipline by producing new
% insights, questions, and conundrums.
% Fertility and reproduction are plausibly a function of both age perspectives,
 % and the question of how to best partition them over time is not investigated here.

The renewal model relates
fertility, mortality, and population structure in a parsimonious form, revealing
the long term consequences of a particular set of conditions. All diverse lines
of contemporary demographic research are relatable to the renewal model, and by
extention to stable population theory. The purpose of this paper is to describe
the Lotka-Leslie renewal model when structured by thanatological age.
%Some might say that stable population
%theory is not relevant to everyday demography, since populations are always
% changing. Even if change precludes the notion of stability as a prediction, models of long-run consequnces still give a
%useful measure of the demographic force implied by a given set of rates.
%In this and other ways, stability is never far away. That population
% projections lead into a stable state means that some form of stability is implied by the routine practice of demography.
%Stability or even stationarity are imaginable for the world population of
% humans one day. Models of stationary populations are easily applied by analogy to
%closed or nearly-closed birth cohorts. Finally, stable population theory is
%relavant for modelling some non-human populations. 
%We have stated that many lifecourse transitions covary more with remaining
%lifetime than with life lived. In an even more general sense, by thinking about
% remaining lifetime as a particular class of time-to-transition, one may generalize the lifetable methods presented here into new area applications, such as education, partnering, the economic lifecycle, or migration.
To present the renewal model based on thanatological age is not necessarily a
suggestion for a better way to model or project population. Rather, the
thanatological renewal model represents an alternative and coherent way to
describe the process of population renewal by relating renewal to a death flow
rather than to a birth flow. I have described a limited case of this
perspective on renewal for the case of stationary populations
\citep{riffe2015force}, and here provide a broader treatment allowing for
growth in the population.
The chronological-age classified renewal model and the thanatological renewal
model both describe the movement of population stocks over time, but from opposite vantage points. %Specifying a model of thanatological renewal expands the toolbox available to approach complex temporal population structure.

The next section of this article describes the lifetable method of transforming age-classified
data into thanatological age classes. I then derive the continuous version of
the thanatological renewal model, and some other stable properties. This is
followed by the matrix version of the model, which is explained alongside the
lifecycle graph. Finally, I explore patterns in stable age structure compared
between the two models. I conclude that thanatological renewal offers
a new perspective on the familiar Lotka-Leslie model.

\section*{Requisites of a thanatological renewal model}
The Lotka-Leslie renewal model \citep{leslie1945use,sharpe1911problem} is
situated in a unisex homogenous population framework. The model expressed in continous mathematics contains a
component for survival, $\ell(a)$, a scaling factor for growth, $e^{-ra}$, and a
component for fertility, $m(a)$:

\begin{equation}
\label{eq:lotka}
1 = \int_{a=0}^\infty l(a)e^{-ra}m(a)\dd a \ep
\end{equation}

Each of these elements is structured by
chronological age, $a$. The model can be approximated by observed data because
chronological age is typically known. The survival function, $\ell(a)$, derives
from a lifetable, usually based on age-specific event-exposure mortality rates,
which requires knowing the age of decedentsand an estimate of the population
exposed to risk in the same age classes. Female fertility rates, $m(a)$, are
classified by age-of-mother

\subsection*{A lifetable approximation of thanatological age}
%Direct information on the timing of demographic transitions with respect to
%death is available retrospectively from linked population registers or
%panel surveys with mortality follow-ups.
%Some surveys also ask respondents in various ways about their perceived future
% lifetime, which may underly those demographic transitions governed to some degree by volition.

In general, to make a lifetable-based projective statement about the remaining
time distribution of a current stock of population, it is best to to account for population subgroups
with different mortality and to account for uncertainty about future changes in
mortality.
The present inquiry approximates thanatological age using lifetable-projected remaining lifetime,
which is intended to approximate remaining lifetime under the strong
assumptions of homogenous populations und fixed vital rates. 
%Chronological age-classified counts are redistributed over future death times
% according to a fixed mortality schedule.
%In general, a lifetable approach to transforming chronologically classified
%data into thanatologically classified data is sound if the given lifetable
%accurately describes the future mortality of the age-classified population, and
%if the chronological and thantological age patterns are independent.
Under these assumptions, each birth cohort undergoes the same
pattern of attrition. In a period setting, this means that we can distribute the
remaining lifespans of the population in each age class using the same
lifetable.
%This is the case we explore, and so the following formulas assume a single lifetable pattern.

@@

A convenient discretization of \eqref{eq:vaupel1} is to simply work with the
$d_a$ column of the lifetable (here with single ages):
\begin{equation}
\label{eq:dyadiscret}
f_{y, a} =  \frac{d_{a+y}+d_{a+y+1}}{L_a + L_{a+1}} \ec
\end{equation}
where $L_a$ is lifetable exposure.\footnote{Equation~\eqref{eq:dyadiscret}
will incur some error depending on the method used to derive $L_a$. To remove
this error, rescale over the $a$ margin so that each distribution conditioned on
$a$ sums to 1. Working with $\ell_a$ rather than $L_a$ eliminates error and
yields similar results, but is not necessarily better.} One produces in this way a probability density function over future death times for each chronological age,
$a$. The $f(y|a)$- weighted average of $y$ for a given $a$ returns the familiar
remaining life expectancy column of the lifetable, $e(a)$. If $P(a)$ is a
chronological age-classified population count, we can derive $P(y)$, a
thanatological age-classified population count using\footnote{A
different-looking but identical definition is found in
\citet{brouard1989mouvements}.}

@@

Lifetable-derived thanatological age structure is
probabilistic until after death, since it refers to the future according to an
assumed mortality pattern; in this case a period mortality pattern is used.
As such, Figure~\ref{fig:USrecomp} is \textit{projective} in nature, a
forward-looking glance at population attrition given the population stock and lifetable of a
particular moment, whereas Figure~\ref{fig:USdecomp} is \textit{reflective} in
nature, since a population's chronological age-structure is mostly the fruit of past
fertility, but also to a lesser degree migration and attrition.

\subsection{Thanatological fertility rates.}
Any count
classified by chronological age can be reclassified in this way given a
pertinent lifetable. To
derive thanatological fertility rates under these assumptions, $\gamma(y)$, apply \eqref{eq:transform} to
chronological age-classified birth counts, $B(a)$, to get $B(y)$ and again to
exposure-to-risk, $E(a)$, to calcualte $E(y)$ and then divide in the usual way:

@@

\FloatBarrier
\section*{The thanatological renewal model}
The renewal model that follows uses thanatological fertility rates, $\gamma(y)$,
the form of which is suggested in Figure~\ref{fig:FySpaghetti}, and we refer to
a homogenous, closed, unisex population based on female vital rates. Assuming constant vital rates, the births for the present year are given by 
\begin{align}
B(t) =& \int _0^\infty P(y,t)\gamma(y) \dd y = \int _0^\infty
P(a,t)m(a) \dd a \ec \intertext{where $a$ indexes chronological age, $y$ indexes
thanatological age, $P(a), P(y)$ are population counts, and $\gamma(y), m(a)$ are
exact fertility probabilities (rates), best imagined
as the fertility rates of females born to females. The thanatological
(first) integral can be broken down back in terms of chronological age:} =&
\int_{y=0}^\infty \int_{a=0}^\infty \gamma(y)P(a,t)f(y|a) \dd a \dd y \ec
\intertext{where $f(y|a)$ is defined per equation~\eqref{eq:vaupel1}, so
$d(a)/\ell(0)$.
We can relate the present population to past births with $P(a,t) = B(t-a)\ell(a)$:}
\label{eq:ergostep1}
=& \int_{y=0}^\infty \int_{a=0}^\infty \gamma(y) B(t-a)d(a+y)\dd a \dd y \ep 
\intertext{Eventually strong ergodicity will assert itself, and $B(t)$ will
be related to $B(t-a)$ according to a constant factor $e^{ra}$, where $r$ is the familiar
intrinsic rate of growth:}
\label{eq:ergostep2}
=& \int_{y=0}^\infty \int_{a=0}^\infty \gamma(y) B(t)e^{-ra}d(a+y)\dd a \dd y
\ep
\intertext{Divide out $B(t)$ to get back a familiar-looking renewal equation:}
\label{eq:thanoren}
1 =& \int_{y=0}^\infty \int_{a=0}^\infty \gamma(y) d(a+y)e^{-ra}\dd a \dd y
\ep
\end{align}
Now compare this to Lotka's chronological formulation in
Equation~\eqref{eq:lotka},which can be expressed as well as:
\begin{equation}
 1 = \int_{a=0}^\infty \int_{y=0}^\infty m(a)d(a+y)e^{-ra}\dd y \dd a \ec
  \label{eq:lotka2}
\end{equation}
and note that \eqref{eq:lotka2} and \eqref{eq:thanoren} are really quite
similar, since $\int _{y=0}^\infty d(a+y)\dd a = l(a)$. For intuition, notice
that $l(a)$ is here decomposed into $d(a)$. Imagine a matrix of this
decomposition, where one margin is chronological age and the other margin is
thanatological age.
For the chronological case \eqref{eq:lotka}, we multiply chronological
age-specific fertility rates, $m(a)$, over the chronological age margin, and for the thanatological case we multiply $\gamma(y)$ over the thanatological age margin.

@@

\section*{Discussion}
The thanatological renewal model is valid, but this perspective may take some
time to gestate before accepting that it is also a sound description of how
populations renew. What should one imagine under the model of thanatological
population renewal? A useful mnemonic bases itself on Figures~\ref{fig:USdecomp}
and \ref{fig:USrecomp}. In the chronological age-structured model, new
generations appear at the bottom of the pyramid, and move up one class per year.
All age-classes are subject to attrition, which is spread out over ages and not
readily visible in the pyramid. In the thanatological \textit{leaf}, each birth cohort increments to the population over the whole range of thanatological age
according to $d(a)$, as seen in \eqref{eq:thanoren}, becoming the
shaded layers seen in Figure~\ref{fig:USrecomp}. Each horizontal step is a death
cohort, and these move one step down the pyramid each year without any decrement
(indeed incrementing due to births) until reaching the very bottom. In short,
the locations of increment and decrement, and the direction of movement (when
so visualized) are all switched. The chronological and thanatological renewal
models are almost perfectly opposite descriptions of the same process.

@@

\section{Equivalence in stability}
\label{app:B}
This appendix contains a proof that the thanatological renewal equation implies
the same intrinsic growth rate, $r$, as the Lotka-Euler model if
the starting population is stable, given a particular set of thanatological
fertility rates.
This is identical to claiming that the right side of the renewal equation is
equal given the same $r$. Since there is a unique real solution, demonstrated in the previous appendix, then it is sufficient to
show that the two integral equations are equal under these conditions. The basic
relationship is:
\begin{align}
\int _{a=0}^\infty \ell(a)e^{-ra}f(a) \dd a &= \int _{y=0}^\infty \int
_{a=0}^\infty d(a+y)e^{-ra}\gamma (y) \dd a \dd y
\intertext{when}
\gamma(y) &= \frac{b^\star(y)}{c^\star(y)} \\
b^\star (y) &= \int_{a=0}^\infty b(a) \mu(a+y)\frac{\ell(a+y)}{\ell(a)} \dd a \\
c^\star (y) &= \int_{a=0}^\infty c(a) \mu(a+y)\frac{\ell(a+y)}{\ell(a)} \dd a
\intertext{Note that:}
c(a) &= \frac{\ell(a)e^{-ra}}{\int \ell(a) e^{-ra} \dd a} \\
b(a) &= m(a)c(a)
\end{align}Replacing $\gamma(y)$ with the full expanded
expression and plugging into the thanatological renewal equation gives:
\begin{align}
=& \int _{y=0}^\infty \int
_{a=0}^\infty d(a+y)e^{-ra} \frac{\frac{\int_{t=0}^\infty
m(t)\ell(t)e^{-rt}\mu(a+y)\frac{\ell(t+y)}{\ell(t)} \dd t}{\int \ell(t) e^{-rt} \dd
t}}{\frac{\int_{t=0}^\infty \ell(t)e^{-rt} \mu(t+y)\frac{\ell(t+y)}{\ell(t)} \dd t}{\int
\ell(t) e^{-rt} \dd t}} \dd a \dd y
\intertext{The denominator of $c(a)$ cancels out:}
=& \int _{y=0}^\infty \int
_{a=0}^\infty d(a+y)e^{-ra} \frac{
\int_{t=0}^\infty
m(t)\ell(t)e^{-rt}\mu(t+y)\frac{\ell(t+y)}{\ell(t)} \dd t}{
\int_{t=0}^\infty \ell(t)e^{-rt} \mu(t+y)\frac{\ell(t+y)}{\ell(t)} \dd t} \dd a \dd y
\intertext{Some $\ell(a)$'s also cancel out:}
=& \int _{y=0}^\infty \int
_{a=0}^\infty d(a+y)e^{-ra} \frac{
\int_{t=0}^\infty
m(t)e^{-rt}\mu(t+y)\ell(t+y) \dd t}{
\int_{t=0}^\infty e^{-rt} \mu(t+y)\ell(t+y) \dd t} \dd a \dd y
\intertext{Now $d(a+y) = \mu (a+y)\ell(a+y)$, so we can cancel the inner integral
with the denominator:}
=& \int _{y=0}^\infty 
\int_{a=0}^\infty
m(a)e^{-ra}\mu(a+y)\ell(a+y) \dd a \dd y
\intertext{Note also that $l(a) = \int _{=a}^\infty d(x) \dd x$ (present
livings are future deaths), which brings us back to the chronological
formulation:} =& \int_{a=0}^\infty e^{-ra}\ell(a)m(a)
\dd a
\end{align}
This finding essentially proves that in the stable population there is at least
one thanatological fertility rate schedule that will satisfy the constraints of
the lifespan distribution and $r$. However, it is not necessary to derive the
thanatological fertility distribution in the way prescribed here, and indeed an
infinite number of such fertility distributions would satisfy the same
constraints. Imagine these fertility rates cross-classified in
both age dimensions, where two marginal sums give the thanatological and
chronological rates, as they would be calculated over the
corresponding cross-classified exposures. The shape of this rate surface can be
shifted around in infinitely many ways that would preserve a given growth rate
and lifespan distribution, and the above proof shows one such mapping.
